---
title: "Feature-based forecasting algorithms for large collections of time series"
author: "George Athanasopoulos, Rob J Hyndman, Pablo Montero-Manso & Thiyanga Thalagala"
date: "22 February 2019"
abstract: "I will discuss two algorithms used in forecasting large collections of diverse time series. Each of these algorithms uses a meta-learning approach with vectors of features computed from  time series to guide the way the forecasts are computed. In FFORMS (Feature-based FORecast Model Selection), we use a random forest classifier to identify the best forecasting method using only time series features. A key advantage of our proposed framework is that the time-consuming process of building a classifier is handled in advance of the forecasting task at hand, and only the selected forecasting model needs to be computed in real time. In FFORMA (Feature-based FORecast Model Averaging), we use gradient boosting to obtain the weights for forecast combinations using as inputs only a vector of time series features. This is slower than FFORMS (because forecasts from all candidate models must be computed), but it provides substantially more accurate forecasts. Both approaches perform very well compared to competitive methods in large forecasting competitions, with FFORMA achieving 2nd place in the recent M4 forecasting competition."
fontsize: 14pt
titlefontsize: 22pt
output:
  binb::monash:
    fig_height: 5
    fig_width: 8
    highlight: tango
    incremental: no
    keep_tex: no
    includes:
      in_header: preamble.tex
colortheme: monashwhite
toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE,
  dev.args = list(bg = grey(0.9), pointsize = 11)
)
library(tidyverse)
library(Mcomp)
library(fpp2)
library(GGally)
library(tsfeatures)
source("functions.R")
set.seed(20180605)
options(digits = 3, width = 63)
```


# Makridakis forecasting competitions

## M competition: 1982

\placefig{0.1}{1.4}{height=8.2cm,width=10cm}{M1}

\only<2->{\placefig{1}{4.3}{height=4cm,width=10cm,keepaspectratio}{SMakridakis}}

\only<3>{\begin{textblock}{5.5}(6.5,2)
  \begin{block}{M-competition}
  \begin{itemize}
  \item 1001 series from demography, industry, economics.
  \item Annual, quarterly, monthly data.
  \item Anyone could submit forecasts.
  \item Multiple forecast measures used.
  \end{itemize}
  \end{block}\end{textblock}}
}

## M3 competition: 2000

\full{M3paper}

## M3 competition: 2000
\fontsize{13}{14}\sf
\begin{block}{}
``The M3-Competition is a final attempt by the authors to settle the accuracy issue of
various time series methods\dots\ The extension involves the inclusion of more methods/ researchers (in particular in the areas of neural networks and expert systems) and more series.''
\end{block}

  * 3003 series
  * All data from business, demography, finance and economics.
  * Series length between 14 and 126.
  * Either non-seasonal, monthly or quarterly.
  * All time series positive.

## M4 competition: 2018

\full{m4}

## M4 competition: 2018

 * January -- May 2018
 * 100,000 time series: yearly, quarterly, monthly, weekly, daily, hourly.
 * Point forecast and prediction intervals assessed.
 * Code must be public
 * 248 registrations, 50 submissions.

\pause

### Winning methods
 1. Hybrid of Recurrent Neural Network and Exponential Smoothing models
 2. FFORMA: Feature-based FORecast Model Averaging (based on FFORMS)


# Time series features

```{r scalem3}
scalem3 <- list()
for (i in 1:3003)
{
  scalem3[[i]] <- M3[[i]]$x - min(M3[[i]]$x)
  scalem3[[i]] <- as.numeric(scalem3[[i]] / max(scalem3[[i]]))
}
```

## Time series features
\placefig{9.1}{1.5}{width=3.6cm}{tukey}
\begin{textblock}{3}(9.7,6.5)\small\textit{John W Tukey}\end{textblock}
\begin{textblock}{8}(0.7,1.2)
\begin{alertblock}{Cognostics}
Computer-produced diagnostics\\ (Tukey and Tukey, 1985).
\end{alertblock}
\end{textblock}\pause
\vspace*{2.5cm}

\alert{Examples for time series}

  * strength of seasonality
  * size and direction of trend 
  * lag correlation
  * spectral entropy

\vspace*{0.3cm}
\begin{block}{}
Called ``features'' in the machine learning literature.
\end{block}


## Feature-space of time series

\fontsize{9}{10}\sf
\begin{textblock}{2.8}(1.5,1.5)
\begin{block}{\fontsize{10}{10}\sf
 STL-decomposition}
$Y_t=T_t+S_t+R_t$
\end{block}
\end{textblock}

\begin{textblock}{8}(5.4,1.3)
\begin{itemize}
\vfill\item strength of trend: max$[0,1-\frac{Var(R_t)}{Var(Y_t-S_t)}]$
\vfill\item strength of seasonality: max$[1-\frac{Var(R_t)}{Va(Y_t-T_t)}] $
\end{itemize}
\end{textblock}

\vspace{2.2cm}


\def\begincols{\begin{columns}}
\def\endcols{\end{columns}}
\def\begincol{\begin{column}}
\def\endcol{\end{column}}


\begincols
\begincol{.48\textwidth}


```{r, warning=FALSE, message=F, echo=FALSE, fig.height=8}
library(Mcomp)
library(ggplot2)
library(grid)
library(gridExtra)
library(ggrepel)
library(png)
library(tsfeatures)
library(tidyverse)
library(ggpubr)

# Extract required series
series_id <- c("N0001", "N0633", "N0625", "N0645","N1912", "N2012")
six_series <- lapply(M3[series_id], function(u){u$x})
p <- lapply(six_series,
function(u) {autoplot(u) + xlab("") + ylab("")}
)
for (i in seq_along(six_series))
p[[i]] <- p[[i]] + ggtitle(series_id[i])+theme(title =element_text(size=25, face='bold'))
grid.arrange(grobs = p, ncol = 2)

```

\endcol \begincol{.48\textwidth}


\pause

```{r fig2, fig.height=8, message=F, echo=F, warning=F}
df <- tsfeatures(six_series, c("stl_features")) %>%
  select(trend, seasonal_strength) %>%
  rename(seasonality = seasonal_strength) %>%
  replace_na(list(seasonality = 0))
df$id <- names(six_series)
ggplot(df, aes(x = trend, y = seasonality)) +
  geom_point(size = 5, color = 'red') +
  xlim(0, 1) + ylim(0, 1) +
  coord_fixed() +
  geom_text_repel(
    aes(label = id),
    colour = "black",
    size = 10,
    box.padding = unit(0.5, "lines")
  ) +
  theme(legend.position = "none")
```

  \endcol
\endcols

\only<3->{
\begin{textblock}{10.4}(1.2,4.2)
\begin{alertblock}{Key idea:}
\fontsize{14}{16}\sf Can we use time series features to guide us in selecting forecasting methods.
\end{alertblock}
\end{textblock}}

# Feature based forecasting algorithms

## Features used to select a forecasting model

\begin{textblock}{12}(0.1,1.3)\small
\begin{multicols}{2}
  \begin{itemize}\tightlist
    \item length
    \item strength of seasonality
    \item strength of trend
    \item linearity
    \item curvature
    \item spikiness
    \item stability
    \item lumpiness
    \item parameter estimates of Holt's linear trend method
    \item spectral entropy
    \item Hurst exponent
    \item nonlinearity
    \item parameter estimates of Holt-Winters' additive method
    \item unit root test statistics
    \item crossing points, flat spots
    \item peaks, troughs
    \item ACF and PACF based features - calculated on raw, differenced, and remainder series.
    \item ARCH/GARCH statistics and ACF of squared series and residuals.
    \end{itemize}
\end{multicols}
\end{textblock}

## Models included

1. Naïve
1. Seasonal naïve
1. Random walk with drift
1. Theta method
1. ARIMA
1. ETS
1. TBATS
1. STL decomposition with AR for seasonally adjusted series

## FFORMS

**F**eature-based **FOR**ecast **M**odel **S**election

\textcolor{red}{Offline}

- A classification algorithm (the meta-learner) is trained.

\textcolor{red}{Online}

- Calculate the features of a time series and use the pre-trained classifier to identify the best forecasting method.


## FFORMS: reference set
 
 \centerline{\includegraphics[width=\paperwidth]{fw4.png}}

## FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{fw5.png}}

## FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{fw6.png}}

## FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{fw7.png}}

## FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{fw8.png}}


## FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{fw9.png}}

## FFORMS: Meta-data

\centerline{\includegraphics[width=\paperwidth]{fw10.png}}

## FFORMS: Random-forest classifier

\centerline{\includegraphics[width=\paperwidth]{fw11.png}}


## FFORMS: Online part of the algorithm

\centerline{\includegraphics[width=\paperwidth]{fw12.png}}


## FFORMS: Online part of the algorithm

\centerline{\includegraphics[width=\paperwidth]{fw13.png}}

## FFORMS: Online part of the algorithm

\centerline{\includegraphics[width=\paperwidth]{fw14.png}}


## \fontsize{15}{15}\bf\sffamily FFORMA: Feature-based FORecast Model Averaging

 * Like FFORMS but using gradient boosted trees (xgboost) rather than random forest.
 * Trained on temporal holdout version of M4 dataset, where size of test sets equal to required forecast horizons
 * Optimization criterion: forecast accuracy not classification accuracy.
 * Probability of each model being best is used to construct model weights for combination forecast.
 * 5 days computing time.


## \fontsize{15}{15}\bf\sffamily FFORMA: Feature-based FORecast Model Averaging

\placefig{0.2}{1.4}{width=12.4cm, trim=0 0 0 70, clip=true}{fforma_graphic}


\vspace*{5.2cm}

### M4 competition results (based on average OWA)

```{r, results='asis'}
tribble(
    ~Place, ~OWA, ~Method,
    "1st", 0.821, NA,
    "2nd", 0.838, "(FFORMA)",
    "3rd", 0.841, NA
  ) %>%
  baretable(digits=3)
```




## \fontsize{15}{15}\bf\sffamily FFORMA: Feature-based FORecast Model Averaging

\only<1>{\placefig{0.2}{1.4}{width=12.4cm, trim=0 0 0 0, clip=true}{prototypes}}
\only<2>{\placefig{0.2}{1.4}{width=12.4cm, trim=0 0 0 0, clip=true}{prototype1}}
\only<3>{\placefig{0.2}{1.4}{width=12.4cm, trim=0 0 0 0, clip=true}{prototype2}}
\only<4>{\placefig{0.2}{1.4}{width=12.4cm, trim=0 0 0 0, clip=true}{prototype3}}
\only<5>{\placefig{0.2}{1.4}{width=12.4cm, trim=0 0 0 0, clip=true}{prototype4}}
\only<6>{\placefig{0.2}{1.4}{width=12.4cm, trim=0 0 0 0, clip=true}{prototype5}}



## Papers and packages
\fontsize{14}{17}\sf

\begin{block}{R packages}
\begin{itemize}\tightlist
 \item \alert{tsfeatures}: Calculating time series features. \newline\url{github.com/robjhyndman/tsfeatures}
 \item \alert{seer}: FFORMS --- selecting forecasting model using features. \newline\url{github.com/thiyangt/seer}
 \item \alert{M4metalearning}: FFORMA -- forecast combinations using features to choose weights. \newline\url{github.com/robjhyndman/M4metalearning}
\end{itemize}
\end{block}

\begin{alertblock}{Papers}
Available from \url{robjhyndman.com}
\end{alertblock}
